{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation

### 2021-01-28 documents/background.tex
Progress on RPi cal
dammit the precursor's at the wrong frequency

Note that it is possible that this has already been worked on. The last review was several months ago, and the literature is changing very rapidly.

@@

@@

Electroporation of viral membranes and capsids has been observed.

Mizuno \cite{Inactivation1990} inactivate SVD Enterovirus (a 27 nm non-enveloped encapsidated ssRNA) and EHV (a 100 nm enveloped dsDNA) with 100\% effectiveness at 120 pulses, about 500 nanoseconds each, at 3 MV/m.

The thermal artifact cannot necessarily be discounted. The energy \footnote{(43 cal/cc * (3 microliters)) / (3 milligrams * 4.2 J/(g K))}. The electrodes were un-insulated, so electrochemical effects are possible.

Interestingly

Madiyar observe rapid permeabilization at a field of approximately $10^7$ V/m, in the Vaccinia poxvirus (a 300 nm enveloped dsDNA virus). \cite{AC2017}. Interestingly, the same effect is not observed in T4 \cite{Manipulation2013}.

@@

This idea, that resonances, is backed up by measurements by Gabriel; the slowest relaxation time in the microwave regime appear to be in the mammary glands, and. Small resonances will be disguised by the aggregation, however.

@@

\begin{fquote}[NATO Research Task Group 189][ \cite{treatyelectromagnetic}]
HFM-189 found no published and replicated adverse health effects or biological mechanisms, beyond
thermal interaction, for pulses shorter than 100 ms which suggested that neither the peak E-field limit in the
IEEE C95.1-2005 safety standard [n.b. chapter 4.3, Table 9, subsection e. peak E field 100 kV/m] nor the proposed limit in the Directive 2004/40/EC [33] (subsequently promulgated as 2013/35/EU [34]) have scientific basis. 
\end{fquote}

" physical laws governing the propagation of E-fields in air already limit the maximum allowable peak E-field at ~3 MV/m (air breakdown). Current research efforts by members to expose biological organism(s), tissues, and cells to environmental
fields up to this magnitude have been unable to elicit an acute biological response."

@@

C95.1-2019, B.4.3 Rationale for pulsed RF field limits

@@

Pakhomov \cite{Comparative} expose frog heart cells, and find no effects on functioning pacemaker cells at 0.9 kV/m (although the field inside the waveguide may be lower).

Note: The author not being biologists or microwave engineers, the trained reader may find that we are too verbose in simple technique and methodology and not detailed in another.

>>>>

Note that it is possible that this has already been worked on. The last review was several months ago, and the literature is changing very rapidly. Reuben

@@

for which our working hypothesis is that there are no clinically useful mechanisms,

It stands to reason, then, that any pulse longer than the relaxation time is not efficiently.

vestiges

Less than 1 nanosecond (where poration and thermal effects start to become significant at these fields, and any previous absorbed power has been dispersed in the solvent) but greater than about 10 picoseconds (covered by existing laser research) and electric fields between 0.1 MV/m (for which biological effects at these timescales are only barely plausible) and 10 MV/m (the production of pulses which strain credulity).

as dt decreases, there's lots of room to add more zeros. a pulse 4 orders of magnitude higher than the maximum field used by practically any existing system.

@@

Scope for this effect can be seen in the following contrast. The statement which echos the concensus of a number of literature reviews by standards organizations:

\begin{fquote}[\textit{NATO Research Task Group 189}, as cited by \textit{C95.1-2019, B.4.3 Rationale for pulsed RF field limits}][ \cite{treatyelectromagnetic}]
	HFM-189 found no published and replicated adverse health effects or biological mechanisms, beyond
	thermal interaction, for pulses shorter than 100 ms which suggested that neither the peak E-field limit in the
	IEEE C95.1-2005 safety standard {\normalfont[NB chapter 4.3, Table 9, subsection e. peak E field 100 kV/m]} nor the proposed limit in the Directive 2004/40/EC [33] (subsequently promulgated as 2013/35/EU [34]) have scientific basis.

	Physical laws governing the propagation of E-fields in air already limit the maximum allowable peak E-field at ~3 MV/m (air breakdown). Current research efforts by members to expose biological organism(s), tissues, and cells to environmental
	fields up to this magnitude have been unable to elicit an acute biological response.
\end{fquote}

\subsection{Precedent}

Electropermeabilization of viral membranes and capsids has been observed.

Mizuno \cite{Inactivation1990} inactivate SVD Enterovirus (a 27 nm non-enveloped protein encapsidated ssRNA) and EHV (a 100 nm enveloped dsDNA) with 100\% effectiveness with 120 pulses, about 500 nanoseconds each, at 3 MV/m. Electron microscopy demonstrates that the mechanism is irreversible damage to the capsid or envelope respectively and release of the core.

\begin{autem}
Extreme care must be taken in interpreting the results of any one study in isolation, especially in this field.

In this observation, the thermal artifact cannot necessarily be discounted. The energy dissipated was sufficient to cause a significant temperature rise if localized hot-spots appeared \footnote{(43 cal/cc * (3 microliters)) / (3 milligrams * 4.2 J/(g K)) = ~50 $\Delta$C, whereas melting temperatures are typically about 90 C for capsids and 50 C for envelopes\cite{Stability1985}}. The electrodes were not insulated, so electrochemical effects are possible but unlikely given the short duration of the pulses. 
\end{autem}

Madiyar \cite{Manipulation2013} observe rapid permeabilization (likely through the slow, conventional electroporation mechanism) at a field of poorly-defined intensity up to about 10 MV/m, in the Vaccinia poxvirus (a 300 nm enveloped dsDNA virus). \cite{AC2017}. Interestingly, the same effect is not observed in T4 bacteriophage.

Moreover, these results could easily be explained by a first-order theoretical. Both AFM data and surmised from osmotic pressure, the force scale required to crack the viral envelopes and capsids is about 1 nanonewton.

de Seze

\pagebreak

(we use the term electropermeabilization)

If we compare in-vitro with in-vitro host cells, the issue becomes less clear.

@@

This idea, that resonances, is backed up by measurements on aggregate tissues by Gabriel; the slowest relaxation time in the microwave regime appear to be in the mammary glands, and. Small resonances will be disguised by the aggregation, however.

@@

@@

Ultrasonic excitation or acoustic pulses have been considered in the literature; we have not analyzed these claims in any depth.

@@

Pakhomov \cite{Comparative} find no non-thermal effects on functioning frog heart pacemaker cells at 0.9 MV/m (although the field inside the tissue was not measured and could have been lower).

### 2021-01-28 documents/brillouin_pulse_math.tex
Progress on RPi cal
dammit the precursor's at the wrong frequency

\pagebreak
\section{The Brillouin precursors}

Uzunoglu \cite{Theoretical2020} mention a handy way to circumvent this lossy-flesh problem: the Brillouin and Sommerfeld precursors, and dispersion in general.

@@

@@

@@

>>>>

@@

\pagebreak
\section{Dispersion and the Brillouin precursors}

Uzunoglu \cite{Theoretical2020} mention a handy way to circumvent this lossy-flesh problem: the Brillouin and Sommerfeld precursors, and pulse dispersion in tissues in general.

Of course, 10x attenuation at 1 MV/m still takes us outside. However, the loss that makes phased-array near-field focusing systems impractical at baseband is no longer a concern. A dispersive phased array with more than 10 elements may allow the complete elimination of attenuation without exceeding the limit field at any point in the body.

@@

\footnote{"precursor" is perhaps a misnomer when single pulses are used rather than trapezoidally-modulated tones - Oughston and Najafabadi call these Brillouin pulses \cite{Optimal2015} \cite{Optimal2017} \cite{Heat2010}.}

@@

The aspect that makes the precursors useful is that they sidestep Bragg's Law decay, decaying as $1/\sqrt{z}$ rather than $e^{-z}$. Note that Bragg's law is still strictly obeyed in the frequency domain (excepting second-order "bleaching" or induced transparency-like effects).

### 2021-01-28 documents/charge.tex
dammit the precursor's at the wrong frequency

Induced polarization

Nuclocapsid

The adsorption of metal ions onto the surface of the genome,

@@

\subsection{Birefringence }

\subsection{Electrorotation }

\subsection{Poisson-Boltzmann }

@@

Yang et al determine that in order to fit their microwave absorption cross-section data, and produce the decrease in infectivity via precisely the mechanism they suggest, the "core" and "shell" of Influenza A should have an effective charge of about $q=1 \times 10^7 e^-$.

The mechanism given is that ions from solution counteract the protein charges to form an "electric double layer". In any case, the best representation is two concentric spheres, with a net charge each.

>>>>

\begin{autem}
Deserves a more detailed treatment.
\end{autem}

The adsorption of metal ions onto the surface of the genome
\begin{itemize}
\item The nucleocapsid. 
\item Induced polarization
		Persson cite a re-arrangement of electrons as a source of drag. Polarization-like effects are not expected. We have not followed up on this.

\item 
\end{itemize}

@@

Birefringence e

It is easy to determine an upper limit.

Božič \cite{How2012} consider an empty capsid.

Donnan equilibrium

@@

Yang et al determine that in order to fit their microwave absorption data, and produce the decrease in infectivity via precisely the mechanism they suggest, the "core" and "shell" of Influenza A should have an effective charge of about $q=1 \times 10^7 e^-$.

The mechanism given is that ions from solution counteract the protein charges to form an electric double layer. In any case, the best representation is two concentric spheres, with a net charge each.

### 2021-01-28 documents/paper.tex
Dialogue
More discussion
Building the thing
dammit the precursor's at the wrong frequency

%\subfile{introduction}

@@

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{eppenwolf_2.jpg}
	\caption{\\ The {\it Eppenwolf} pulsed X-Band microwave spectrometer used in this study, pictured with the near-field cuvette installed.}

\end{figure}

@@

\paragraph{\textbf{Subharmonics}}\

@@

@@

@@

@@

>>>>

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{pulse_exposure_setup.JPG}
	\caption{\\ Pulse exposure jig. Autosampler Pulse shaper not shown. Aluminum disc belongs to the scan-conversion digitizer, not used in this study.}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{eppenwolf_2.jpg}
	\caption{\\ The misguided {\it Eppenwolf} pulsed X-Band microwave spectrometer used in this study, pictured with the near-field cuvette installed.}
\end{figure}

@@

@@

@@

\subfile{brillouin_pulse_math}

@@

\subfile{charge}

@@

\subfile{molecular_dynamics}

@@

\subfile{fluorescence}

\subfile{detection}

### 2021-01-28 documents/pulse_generator.tex
dammit the precursor's at the wrong frequency

%!TeX root = brillouin_pulse_math
\documentclass[paper.tex]{subfiles}
\begin{document}

The extremely high losses from such thin coaxial cables should be considered.

\pagebreak
\section{Dispersion and the Brillouin precursors}

Uzunoglu \cite{Theoretical2020} mention a handy way to circumvent this lossy-flesh problem: the Brillouin and Sommerfeld precursors, and pulse dispersion in tissues in general.

Of course, 10x attenuation at 1 MV/m still takes us outside. However, the loss that makes phased-array near-field focusing systems impractical at baseband is no longer a concern. A dispersive phased array with more than 10 elements may allow the complete elimination of attenuation without exceeding the limit field at any point in the body.

The original text of these papers are in German; translations can be found in Léon Brillouin's 1960 
book\cite{Wave1960} on the topic. \footnotemark

These are fascinating structures which arise as a result of two aspects (Albanese 
\cite{Shortrisetime1989}):

\footnote{"precursor" is perhaps a misnomer when single pulses are used rather than trapezoidally-modulated tones - Oughston and Najafabadi call these Brillouin pulses \cite{Optimal2015} \cite{Optimal2017} \cite{Heat2010}.}

\begin{itemize}
	\item The highly dispersive nature of tissue, which causes a peculiar distortion of short pulses,
	\item The variation in loss (the imaginary part of the complex permittivity) and penetration depth versus frequency. 
\end{itemize}

A third aspect leads to the formation of the prototypical Brillouin precursor specifically: any sharp 
beginning or end of a tone, or other sharply defined structure in a waveform, includes harmonics beyond the tone's carrier wave up to the rise time \footnotemark. \wikinote{add image}

The aspect that makes the precursors useful is that they sidestep Bragg's Law decay, decaying as $1/\sqrt{z}$ rather than $e^{-z}$. Note that Bragg's law is still strictly obeyed in the frequency domain (excepting second-order "bleaching" or induced transparency-like effects).

Precursors have been used in practice, by Ong \cite{Detection2003} using simple square waves to detect respiration\footnotemark, Ossberger \cite{Noninvasive2004} for the same purpose; and various techniques for breast cancer detection.

Also another really good paper that I can't find.

\footnotetext{Well, Ong don't really observe the precursor itself, it seems.}

By putting more power in higher frequencies,

While it is clearly possible to produce this abnormal penetration (any sharp edge suffices), it was not obvious to us that a waveform 
could be constructed that would be an effective driver for the normal modes in question. 
Oughston ; if the pulses are too close together, they distort the precursor thus formed.

\footnotetext{it is interesting to note that a similar - if not equivalent - effect leads to the Heisenberg uncertainty principle. The more localized a }

At a depth of 8 cm in Cole-Cole muscle, an 8 GHz resonance, attack pattern alpha appears to achieve 4 picometers of oscillation amplitude, rather than 0.01 picometers with $\sin(t\ \omega_{resonance})$. The sharp edges are at approx. 250 GHz.

At 9 GHz, assuming that a torso is about 10 penetration depths in radius and the Brillouin train 
drives the virus with an equally, implementing this would decrease the field at the skin required 
for complete eradication of the virus from the body to a perfectly practicable 1 kV/m, rather than 
an air-ionizing 6.6 MV/m ($\frac{1}{\sqrt{10}}=0.316$ versus $e^{-10}=4.5e^{-5}$).

However, the energy contained in such a standard pulsed-sine or gaussian train decays very quickly,

%http://web.physics.ucsb.edu/~fratus/phys103/LN/DHM.pdf
% section on Greens functions
% and also 
% and
% https://www.int.washington.edu/users/dbkaplan/228_01/green.pdf

Note: this isn't truly transient - the fourier integral is still periodic, wrapping around every x cycles. It's just more transient than the steady-state.

attack pattern alpha

Unfortunately, despite NLTLs producing strong wideband components, in cursory simulations they do 
not appear to satisfy the requirements for Brillouin pulses. Our working hypothesis is that this is 
because of destructive interference from the falling edge that immediately follows; but this is 
mere speculation.
\begin{autem}
	This should be replicated.
\end{autem}

temporal Soliton-producing nonlinear transmission lines, and NLTL oscillators constructed therefrom, particularly well suited

Transient solution

\pagebreak
A qualitative statement of the problem:

\begin{toolchain}
An arbitrary waveform propagates through a dispersive, lossy medium with a certain analytic complex 
permittivity.

It then drives a damped harmonic oscillator that is initially at rest.\\

What waveform produces the peak transient amplitude on the oscillator?
\end{toolchain}

\subsection{Analytic techniques}

\paragraph{\textbf{Subharmonics}}\

\begin{fquote}[Wait][ \cite{Propagation1965}]
	In such cases, it may be feasible to evaluate the integral by a purely numerical procedure. With the wide availability of the digital computer, this is certainly fashionable at the moment. Consequently, one might say that the problem has been solved and no further discussion is needed. \\

	However, it would be a pity if one accepted this answer since all physical insight into the nature of transient processes has been ignored. 
\end{fquote}

\begin{autem}
Reflection from tissue interfaces?

Thermal noise will kick the oscillator around; it won't necessarily start at rest. Will this affect the 
\end{autem}

One might suggest try to find the optimal transient solution first, and then try to build an input waveform deconvolved through the  transfer function of the tissue. This hardly seems liable to produce a truly optimal solution, however.

Each problem has been solved separately. There are some nearly equivalent but not mathematically identical formulations for optimal control of quantum systems with dispersion.\footnotemark However, in the classical domain, this appears to be unprecedented in the literature.

\footnotetext{While not strictly related, \cite{Wavepacket1989} tune a light pulse to match the wavefunction of the desired chemical products in a photo-reaction, which is just pretty darned awesome}

Also, this seems similar to a pulse propagating through the tissue's Cole medium and then dissipating the maximum power in a uniform Lorentz medium, which might be another route to attacking it.

Another route might be to set up two ODEs for the wave equation and the oscillator and solve them simultaneously.

Progress was hindered by the ~6 conventions for Fourier normalization, the sign or handedness of the imaginary number. What self-respecting field has completely antipodal definitions of the same concept.

oh, on the same topic, can we finally settle the i / j convention. Also, everyone that doesn't explicitly note

Optimal pulses for several situations in Debye and Lorentz media have been determined by Oughstun's asymptotic method, which appears to be highly satisfactory in terms of physical intution, and also avoids truncation issues that befall fourier-transform techniques. In general, these optimal pulses are simple gaussian monopulses or higher-order gaussian derivatives \cite{Optimal2017} \cite{Optimal2015}.

Macke \cite{Simple2012} have come up with a number of other analytic dispersion representations for various signals, such as chirped gaussian pulses.

%\printendnotes

\subsubsection{Zhu et al 's dispersive waveform optimization propagation formulation}

The most promising analytical technique that we explored was essentially equivalent to Zhu, Hum, 
Costas formulation in \cite{Microwave2012b}, previously discussed by Pozar \cite{Waveform2003}, 
This is based on a lagrange multiplier technique and variational 
calculus\cite{Methods1989} \footnotetext{Based on personal communication with Professor Costas Sarris}.

$$ e_x(z,t) = \frac{1}{\sqrt{2 \pi}} \int_{-\infty}^{+\infty}{F(\omega) e^{- j (\omega/c_0)n(\omega)z}\ e^{j\omega t} d\omega} $$

where $\int e^{j \omega t} d\omega$ is the inverse fourier transform.

$$W = \frac{1}{\eta_0} \int_{-\infty}^{+\infty}{(n_r(\omega))\ |F(\omega)|^2}\ d\omega$$\\

Where $n(\omega)$ is a complex refractive index per Debye relaxation equation, $n_r()$ is the real part of the same, everything else is a real constant. Set up a functional with a lagrange multiplier\\

$$\xi = -e_x(z,t) + \lambda W$$\\

Take the functional ("variational") gradient with respect to $F(\omega)$.

$$ \nabla \xi = -\frac{1}{\sqrt{2\pi}} \left(e^{- j (\omega/c_0)n(\omega)z}\right)^\star \  e^{-j\omega T} + \lambda...$$

At this point, you could choose to perform a second integral over the time-domain Greens function solution to the harmonic oscillator\cite{Complex2020}, eq. 4 and 15,

$$x(t) = \int_{-\infty}^{+\infty}{G(t,t')\ \frac{q\cdot e_x(z,t'))}{m}\ } dt'$$\\

(eq. 15, for the underdamped case)

$$G(t,t') = UnitStep(t-t') \frac{1}{\sqrt{\gamma^2-\omega_0^2}  }\sinh\left({\sqrt{\gamma^2-\omega_0^2}}\ (t-t')\right)$$

But a much more straightforward

Where * is the complex conjugate.\\

The $\nabla H() = H()^\star$ thing seems to come from Franks, Signal Theory, 1969, page 140 onwards; I cannot make heads or tails of his formulation.

Neither Maxima, SageMath, Mathematica's VariationalD, Matlab's Fundiff, nor Maple's 
functionalDerivative seem to yield any meaningful result for grad xi.\\

Then set

$$x(t) = \int_{-\infty}^{+\infty}{G(t,t')\ \frac{q\cdot e_x(z,t'))}{m}\ } dt'$$\\

Anyhow, then modify the lagrange multiplier,

$$\xi = -x(t) + \lambda W$$\\

And now,

$$\nabla \xi = 0$$

Pulse energy, rather than peak field, is not obviously the ideal constraint for this problem; however, it is the one most compatible with lagrange multiplier techniques.

Tufts \cite{Optimum1964}

The Euler-Lagrange equation is usually brought up in the case of the Brachistochrone problem, or 
the Principle of Least Action (see Feynman lectures). In these types of problems, it appears that 
the path of integration through space from endpoints a to b is replaced with the whole frequency 
domain from -$\infty$ to +$\infty$.

Search terms: "first variation", gateaux derivative,

As noted by Pozar, the complex conjugate arises from the concept of "adjointness". Franks.

Note the ni specifically in the complex conjugate!

\subsection{Laziness prevails; a return to numerical optimization}

Numerically solving optimal control problems using CG was noted by Lasdon \cite{conjugate1967} and 
revisited by Kek \cite{Conjugate}. They reduce the computational workload by first computing the 
gradient function by hand.

Using an energy constraint as above appears to tend towards a square step function or a sharp impulse, perhaps as a result of Pontryagin's maximum principle\cite{Optimum1964}.

However, a maximum absolute value constraint produces fantastic results.

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{muscle_pulse_8cm_minimize_test1}
	\caption{}
\end{figure}

The (noisy) output of the optimization for z=0.08 m in muscle. The low-frequency component is believed to be due to the FFT window size. Increasing the window size with the same sample rate caused CG convergence to stall. However, it's interesting that this appears to result in a larger amplitude.

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{muscle_pulse_8cm_minimize_test2}
	\caption{Second.}
\end{figure}

>>>>

%!TeX root = pulse_generator
\documentclass[paper.tex]{subfiles}
\begin{document}
